%\thesistitlepage % make the ordinary titlepage
\hypersetup{pageanchor=false}
%\include{summary}

\chapter*{Preface}
The following thesis deals with rule-based event correlation alerting in real time, based on Windows Event Log. This is a a Master's thesis in MIS (Master in Information Security) at NTNU Gjøvik, carried out during the spring semester in 2020. The idea for the thesis was brought up during my work at BDO AS, where their CERT was exploring the same issue with creating high fidelity alerts from Windows Event Log in a real-time fashion.

The thesis is targeted at audience from the field of information technology and especially those working with security monitoring. However, no expert knowledge is required as the thesis explain all terms and methods used.\\[2cm]
\iffalse
The following work deals with multinomial malware classification of ten malware families with machine learning algorithms. It is carried out in the context of a Master’s thesis in MIS (Information
Security) at NTNU. It was conducted during the spring semester 2019. The broad idea of the topic
was brought up by my supervisors before it was specified in more detail and finalised in a direct
discussion between us. It is still a relevant topic today since the malware landscape is constantly
growing and evolving and therefore, further research has to be conducted in the area of this topic.
Nonetheless, the thesis is targeted at an audience from the field of information technology with the
focus on forensics. However, no expert knowledge of the malware landscape is needed to understand the elaboration since all used terms and methods are explained.

Here, you give a brief introduction to your work. What it is (e.g., a Master's thesis in AIMT at NTNU, when it was carried out (e.g., during the autumn semester of 2021). If the project has been carried out with a company, you should mention this and also describe the cooperation with the company. You may also describe how the idea to the project was brought up.

You should also specify the assumed background of the readers of this report (who are you writing for).\\[2cm]
\fi
%\begin{center}
%\thesiscampus, 
\thesisdate \\[1pc]
\\[1pc]
%\thesisauthor
%\end{center}

\chapter*{Acknowledgment}
Foremost, I would like to express my sincere gratitude to my supervisor, Ass. Prof. Geir Olav Dyrkolbotn for providing excellent guidance, assistance and support during this thesis. I especially appreciate how my supervisor has facilitated guidance from me as a distance student at NTNU.

A big thank goes to my employer BDO AS, and especially Ingunn Holte and Håkon Lønmo, for allowing me time to research, study and write my thesis while working full-time.

I highly appreciate all motivation and support from friends and family throughout my studies. I especially couldn't have done this without my partner Cecilie which has supported me through all ups and downs along the way. I would also like to mention our son Arthur who never failed to cheer me up when I was stuck or met an obstacle during my writing of this thesis.

\begin{flushright}
M.I.\\[1pc]
\end{flushright}

\chapter*{Abstract}
\iffalse
LES "Scrutiny of the abstract" referert i Scrutiny of the introduction - http://sepwww.stanford.edu/sep/prof/Intro.html
\fi
This is a summary of the thesis including the conclusion of what was discovered.

\chapter*{Sammendrag}
\iffalse
LES "Scrutiny of the abstract" referert i Scrutiny of the introduction - http://sepwww.stanford.edu/sep/prof/Intro.html
\fi
Dette er et sammendrag...

\hypersetup{pageanchor=false}